---
title: "Assignment 1 - Loading Data into a DataFrame"
author: "Nick Climaco"
date: "2023-01-28"
output: 
    html_document: default
    pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Summary : 
The article "Club Soccer Prediction" is a forecasting model of club soccer around the world. 
This article uses a revised version of ESPN's SPI (Soccer Power Index) rating from seasons 2016 to 2022. 
The article shows the probability of a team winning, losing, or drawing their respective games.

Link to the articles: 

https://projects.fivethirtyeight.com/soccer-predictions/

https://fivethirtyeight.com/methodology/how-our-club-soccer-predictions-work/

```{r}
library(dplyr)
# added csv file to my github
data <- read.csv("https://raw.githubusercontent.com/Nick-Climaco/Rdataset/main/spi_matches.csv")
head(data)
```
```{r}
# create a subset from data where the probability of a tie is greater than the probability of either team winning. 
df <- data %>% 
    select(season, league, team1, team2, spi1, spi2, prob1, prob2, probtie, xg1, xg2) %>% 
    rename(expected_goal1 = "xg1", expected_goal2 = "xg2") %>%
    filter(probtie > prob1 & probtie > prob2)
head(df)
```

## Conclusion and Findings :
A very interesting fact based on the data provided, we observe how low the chances are where a tie game is the most probable outcome. 
For further analysis, we could analyze how SPI for each team can impact their likelihood of a match regardless if they are the home or away team. 
It also be interesting to cross-examine the teams with the best SPI across different leagues and even different regions of the world.

