---
title: "Project 1 Chess Rating"
author: "Nick Climaco"
date: "2023-02-15"
class: "DATA 607"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE,tidy.opts=list(width.cutoff=90),tidy=TRUE)
```

# Project Description

In this project, you’re given a text file with chess tournament results where the information has some structure. Your
job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database)
with the following information for all of the players:
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605

# Import Required Packages
```{r}
library(tidyverse)
library(knitr)
```

# Loading and Reading the Data
```{r}
tournament_data <- read.csv("https://raw.githubusercontent.com/Nick-Climaco/DATA-607-NC/main/tournament_info.txt", header = TRUE, skip = 3)
tournament_data2 <- read_lines("https://raw.githubusercontent.com/Nick-Climaco/DATA-607-NC/main/tournament_info.txt")
```

# Extracting the Data from the txt file

### Extraction and Cleaning

```{r}
# matches a string of characters for the first and last names or match first, middle and last names
# simplify =TRUE returns a character matrix instead list of character vector
player_names <- tournament_data %>% 
    str_extract_all("\\w+\\s\\w+ | \\w+\\s\\w+\\s\\w+", simplify = TRUE) %>%
    str_flatten(collapse = ",") %>% 
    str_replace_all(",{2,}" , ",") %>%
    str_remove_all("^$") %>%
    str_split(",") 

player_names <- player_names[[1]] %>%
  str_trim()  # remove leading/trailing white space

player_state <-  tournament_data %>% 
    str_extract_all("\\s([A-Z][A-Z])\\s\\|", simplify =  TRUE) %>%
    str_flatten(collapse = ",") %>%
    str_replace_all(",{2,}" , ",") %>%
    str_remove_all("\\|") %>%
    str_split(",") 

player_state <- player_state[[1]] %>%
  str_trim()

total_number_points <- tournament_data %>%
    str_extract_all("\\d\\.\\d", simplify = TRUE) %>%
    str_flatten(collapse = ",") %>%
    str_replace_all(",{2,}" , ",") %>%
    str_split(",")

total_number_points <- total_number_points[[1]] %>% 
    str_trim()

player_pre_rating <- tournament_data %>%
    # not all the pre rating are the same format
    str_extract_all("R:\\s*(\\d\\d\\d\\d)|R:\\s*\\d\\d\\d", simplify = TRUE) %>%
    str_flatten(collapse = ",") %>%
    str_replace_all(",{2,}" , ",") %>%
    str_remove_all("R: ") %>%
    str_split(",") 

player_pre_rating <- player_pre_rating[[1]] %>%
    str_trim()
```
Adding the cleaned into a data frame...
```{r}
# checking if it worked
df_tournament <- data.frame(Number = seq(1:64),
                            Name = player_names, 
                            State = player_state, 
                            Total_Points = as.numeric(total_number_points), 
                            Pre_Rating = as.numeric(player_pre_rating))


head(df_tournament)
```
#### Create a List of opponents player numbers
```{r}
opponents_data <- tournament_data2[seq(5, 196, 3)]
opponent_num <- str_extract_all(opponents_data, "(?<=\\|([A-Z])\\s{2,3})\\d{1,2}|(?<=\\|)(([A-Z])\\s{4}|\\s{5})") %>%
  unlist() %>%
  as.numeric()

#KEEP TRACK OF REGEX MEANINGS:

#(?<=\\|([A-Z])\\s{2,3}) :regex using a positive look behind token  "(?<=..)", match a capital letter A-Z; whitespace 2-3 times

#\\d{1,2} : matches decimal digit 1-2 times 

#"|" : alternate_match

#(?<=\\|)(([A-Z])\\s{4}|\\s{5}): positive look behind \\|; alternate_match a single capital letter followed by 4 with space or no letter just 5 white spaces 

```

```{r}
# Split the opponent numbers into groups of 7
opponent_groups <- split(opponent_num, rep(1:(length(opponent_num) %/% 7), each = 7))

# Create a matrix to store the total and average pre-ratings
opponent_pre_rating <- matrix(0, ncol = 2, nrow = length(opponent_groups))
colnames(opponent_pre_rating) <- c("Total", "Average")

# Calculate the total and average pre-ratings for each group of opponents
opponent_pre_rating[, 1] <- unlist(lapply(opponent_groups, function(opponents) {
  total_pre_rating <- sum(df_tournament$Pre_Rating[df_tournament$Number %in% opponents], na.rm = TRUE)
  return(total_pre_rating)
}))

opponent_pre_rating[, 2] <- unlist(lapply(opponent_groups, function(opponents) {
  average_pre_rating <- mean(df_tournament$Pre_Rating[df_tournament$Number %in% opponents], na.rm = TRUE)
  return(average_pre_rating)
}))
```
```{r}
df_tournament$Average_Opponents_Rating <- round(opponent_pre_rating[,2], digits = 0)

kable(df_tournament, row.names = NA)
```
## Write csv file
```{r}
write.csv(df_tournament, "chess_tournament.csv", row.names = FALSE)
```

