---
title: 'Basic Report'
author: "Nick Climaco"
date: "`r Sys.Date()`"
output:
    html_document:
        highlight: pygments
        theme: cerulean
        toc: true
        toc_float: true
editor_options:
    chunk_output_type: console
---

# Overview 

With the attached data file, build and visualize  eigenimagery that accounts for 80% of the variability.  Provide full R code and discussion

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```

```{r,echo=FALSE}
library(magick)
library(purrr)
library(gridExtra)
library(imager)
library(raster)
library(jpeg)
library(OpenImageR)
library(ggplot2)
```

# Load JPEG Files
```{r}
path = "~/DataScience/DATA_605_F23/jpg"

files=list.files(path ,pattern="\\.jpg", full.names = TRUE)
num = length(files)
height=1200
width=2500
scale=20
new_height <- height/scale
new_width <- width/scale
```

# Initialize Array
```{r}
array_a <- array(rep(0, num * new_height * new_width * 3), 
                 dim = c(num, new_height, new_width, 3))

for (i in 1:num) {
    temp <- resizeImage(readJPEG(files[i]),new_height, new_width)
    array_a[i, , , ] <- array(temp, dim = c(1, new_height, new_width, 3))
}
```

In this code block, we take a look at each shoe and then for each shoe separates the rgb values and places the values in the matrix flat. 

```{r}
flat <- matrix(0, num, prod(dim(array_a)))

for (i in 1:num) {
    new_array_a <- readJPEG(files[i])
    r = as.vector(array_a[i,,,1])
    g = as.vector(array_a[i,,,2])
    b = as.vector(array_a[i,,,3])
    flat[i,] <- t(c(r,g,b))
}
shoes <- as.data.frame(t(flat))
```

# View the Shoes

```{r}
# plot function from prof's code
plot_jpeg = function(path, add=FALSE) { 
    jpg = readJPEG(path, native=TRUE) 
    res = dim(jpg)[2:1] 
    if (!add) {
      plot(1,1,
         xlim=c(1,res[1]),
         ylim=c(1,res[2]),
         asp=1,type='n',
         xaxs='i',
         yaxs='i',
         xaxt='n',
         yaxt='n',
         xlab='',
         ylab='',
         bty='n'
         )
    }
  rasterImage(jpg,1,1,res[1],res[2])
}

par(mfrow=c(3,3))
par(mai=c(.3,.3,.3,.3))
for (i in 1:num){  
    plot_jpeg(writeJPEG(array_a[i,,,]))
}
```



```{r}
scaled <- scale(shoes, center = TRUE, scale = TRUE)
mean.shoe <- attr(scaled, "scaled:center")
std.shoe <- attr(scaled, "scaled:scale")
```

```{r}
Sigma_ <- cor(scaled)
```

Calculating the eigenvalues and eigenvectors 
```{r}
myeigen <- eigen(Sigma_)

eigenvalues <- myeigen$values
eigenvalues
```

```{r}
eigenvectors <- myeigen$vectors
head(eigenvectors)
```

Compare the original image to the eigenshoe.

```{r}
testing_img <- readJPEG(files[4])

plot(1:2, type = "n", main = "")
rasterImage(testing_img,1,1,2,2)
```

```{r}
scaling <- diag(myeigen$values[1:5]^(-1/2)) / (sqrt(nrow(scaled)-1))
eigenshoes <- scaled%*%myeigen$vectors[,1:5]%*%scaling
par(mfrow=c(2,3))
imageShow(array(eigenshoes[,4], c(60,125,3)))
```

Here, we calculate the cumulative variance explained
```{r}
cumulative_variance <-cumsum(myeigen$values) / sum(myeigen$values)
cumulative_variance
```

Next is to figure out the number of eigenshoes that accounts 80% of the variability

```{r}
num_comp <- which(cumulative_variance >= 0.8)[1]
num_comp
```

```{r}
ggplot(as.data.frame(cumulative_variance), aes(x =1:num, cumulative_variance)) +
    geom_line() + geom_point() +
    labs(
        x = "Number of eigenshoes",
        y = "Cumulative Variance"
    ) +
    scale_x_continuous(breaks = seq(1,17, by = 2)) +
    theme_minimal()
```

Now, we can visually see that after 3 eigenshoes accounts for 80% of the variability

## Resources

[Professor Fulton's Rpubs](https://rpubs.com/R-Minator/eigenshoes)

[Eigenfaces for Dummies](http://jmcspot.com/Eigenface/)

[RDocumentation: OpenImageR](https://www.rdocumentation.org/packages/OpenImageR/versions/1.3.0/topics/resizeImage)