---
title: 'Data 605: Assignment 2'
author: "Nick Climaco"
date: "`r Sys.Date()`"
output:
    html_document:
        highlight: pygments
        theme: cerulean
        toc: true
        toc_float: true
        number_sections: true
editor_options:
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```

## Pset 1 {-}

### Question 1 {-}

Show that $A^TA \neq AA^T$ in general. (Proof and Demonstration)

#### Proof: {-}

Suppose, we have 2x2 matrix A where a,b,c,d$\in$ **R** such that they are distinct

$$A = \begin{bmatrix}
a&b \\
c&d \\
\end{bmatrix} $$

Then, matrix A transpose is 

$$A^T = \begin{bmatrix}
a&c \\
b&d\\
\end{bmatrix} $$

Next, we have to compute $A^TA$ and $AA^T$

$$\begin{align}A^TA = 
\begin{bmatrix}
a&c \\
b&d\\
\end{bmatrix} 
\begin{bmatrix}
a&b \\
c&d\\
\end{bmatrix} \\
\Rightarrow 
\begin{bmatrix} 
a^2+c^2&ab+cd\\
ba+dc&b^2+d^2
\end{bmatrix}
\end{align}$$
and 

$$\begin{align}AA^T = 
\begin{bmatrix}
a&b \\
c&d\\
\end{bmatrix} 
\begin{bmatrix}
a&c \\
b&d\\
\end{bmatrix} \\
\Rightarrow 
\begin{bmatrix} 
a^2+b^2&ac+bd\\
ca+db&c^2+d^2
\end{bmatrix}
\end{align}$$

So,

$$A^TA \neq AA^T \\
\begin{bmatrix} 
a^2+c^2&ab+cd\\
ba+dc&b^2+d^2
\end{bmatrix} \neq 
\begin{bmatrix} 
a^2+b^2&ac+bd\\
ca+db&c^2+d^2
\end{bmatrix} $$

$\square$

#### Demonstration: {-}

Let a=1, b=2, c=3, and d=4

$$A = \begin{bmatrix}
1&2 \\
3&4 \\
\end{bmatrix} and \  A^T = \begin{bmatrix}
1&3 \\
2&4 \\
\end{bmatrix} $$

Then we have,

$$A^TA \neq AA^T \\
\begin{bmatrix} 
10 & 14 \\
14 & 20
\end{bmatrix} \neq 
\begin{bmatrix} 
5 & 11 \\
11 & 25
\end{bmatrix}$$

### Question 2 {-}

For a special type of square matrix A, we get $A^TA = AA^T$. Under what conditions could this be true? 

#### Response {-}

By **Definiton SYM**(Symmetric Matrix) and **Theorem SMS**(Symmetric Matrices are Square), this statement is true if and only if the square matrix A is symmetric. 

where,

$$A = A^T$$

Then,

$$A^TA = AA^T \Leftrightarrow AA^T = A^TA$$

Thus, this statement is true iff matrix A is symmetric.

## Pset 2 {-}

Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.

```{r}

lu_factor <- function(A) { # input is some nxn (square) matrix A, should work on any size matrix
    n <- nrow(A) # get the size of matrix A
    
    lower <- diag(1, n) # SETUP the lower triangle
    upper <- A # SETUP the upper triangle
    
    # start loop at row 2 through lower n rows
    for(i in 2:n){ 
        
        # loop column 1 until i-1 column
        for(j in 1:(i-1)){ 
            
            # find the multiplier needed to eliminate the element in A[i,j]
            row_mupltiplier <- -(upper[i,j] / upper[j,j]) 
            
            # update the entire row with multiplier
            upper[i, ] <- row_mupltiplier * upper[j, ] + upper[i, ] 
            
            # place the negative of the multiplier  to the element in lower[i,j]
            lower[i,j] <- -row_mupltiplier 
        }
    }
    
    
  return(list(Lower_Matrix = lower, Upper_Matrix = upper))
}
   # output is the lower and upper matrices
```

### Test and Check {-}

```{r}
library(matrixcalc)
```

```{r}
B <- diag(1,4)
C <- matrix(c(1,-2,-2,-3,3,-9,0,-9,-1,2,4,7,-3,-6,26,2), nrow=4,ncol=4)
D <- matrix(c(1,2,3,4,5,6,7,8,9),nrow = 3, ncol =3)
```

#### Test B {-}
```{r}
lu_factor(B)
lu.decomposition(B)
```

#### Test C {-}
```{r}
lu_factor(C)
lu.decomposition(C)
```

#### Test D {-}
```{r}
lu_factor(D)
lu.decomposition(D)
```

## Resources {-}
* [RDocument: lu.decomposition() ](https://www.rdocumentation.org/packages/matrixcalc/versions/1.0-6/topics/lu.decomposition)

* [Youtube: LU Decomposition - Shortcut Method](https://www.youtube.com/watch?v=UlWcofkUDDU)
