---
title: 'Data 605: Homework 11'
author: "Nick Climaco"
date: "`r Sys.Date()`"
output:
    html_document:
        highlight: pygments
        theme: cerulean
        toc: true
        toc_float: true
editor_options:
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```

# Overview

> Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)

# Data Exporation 

```{r}
head(cars)
```

```{r}
summary(cars)
```

# Model Building

```{r}
lm_mod <- lm(speed ~ dist, data = cars)
```



## Visualization 

```{r}
plot(cars$dist, cars$speed, xlab = "Stopping Distance", ylab = "Speed")
abline(lm_mod)
```

## Evaluation

```{r}
summary(lm_mod)
```

So, we have a formula 

$$
y = 8.28 + 0.16(dist)
$$

This simple linear model has R-squared of 0.65 meaning the model explains 65% of the data. Looking at the residual statistics, it has seemingly center around around and the iqr is even. 

## Residual Analysis

```{r}
plot(lm_mod$fitted.values, lm_mod$residuals, xlab = "Fitted", ylab = "Residuals")
abline(h =0)
```

```{r}
qqnorm(lm_mod$residuals)
qqline(lm_mod$residuals)
```

The qq plot tells us that the most of the residual follow the theoretical normal where we see some of the points on top deviate from the line.

```{r}
par(mfrow = c(2,2))
plot(lm_mod)

```

