---
title: 'Data 605: Homework 7'
author: "Nick Climaco"
date: "`r Sys.Date()`"
output:
    html_document:
        highlight: pygments
        theme: cerulean
        toc: true
        toc_float: true
editor_options:
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```

# Question 1

> Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y

### Response

$$
\begin{align}
F_Y(y) = P(Y \leq y) &= 1 - P(Y>y) & \text{complement of max} \\
&= 1 - P(min(X_1,X_2 ... X_n) > y) & \text{minimum of the X's} \\
&= 1 - P(X_1 > y, X_2>y, ..., X_n > y) \\
&= 1 - P(X_1 >y) P(X_2>y)...P(X_n > y) & \text{due to mutually independent} \\
&= 1 - P(X_i >y)^n & \text{i.i.d} \\ 
&= 1 - (\frac{k-y}{k})^n &  \\
\end{align}
$$

k-y is the number of remaining outcomes divided by the total number of outcome which in this case is k.

This gives us the  cumulative probability of Y less than or equal y. I think the question wants the probability of Y = y which is the probability density function. So, to remove the lower section of the cdf. We can substract $P(Y \leq y -1 )$ from $P(Y \leq y)$. 

$$
P(Y \leq y -1) =1 - \frac{(k-y+1)^n}{k^n}
$$

So, the distribution function is 

$$
\begin{align}
f_Y(y) &= P(Y = y) \\
&= P(Y \leq y) - P(Y \leq y-1) \\
&= 1 - (\frac{k-y}{k})^n - (1 - \frac{(k-y+1)^n}{k^n}) \\
&= \frac{(k-y+1)^n - (k-y)^n}{k^n}
\end{align}
$$

# Question 2 {.tabset}

>  Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).  This machine has a manufacturer’s expected lifetime of 10 years.  This means that we expect one failure every ten years.  (Include the probability statements and R Code for each part.). 

## 2.A

> What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as a geometric.  (Hint: the probability is equivalent to not failing during the first 8 years..) 

$$
X \sim Geom(\frac{1}{10}) \\ 
P(X > n) = 1 - P(X \leq 8) = 1 - \sum_{i=1}^{8} (0.9)^{i-1}(0.1) \\
\Rightarrow 1 - 0.1(1 + 0.9 + 0.9^2 + ... + 0.9^7) \\
E[X] = \frac{1}{1/10} = 10 \\
\sigma = \sqrt{Var(X)} = \sqrt{90} = 3\sqrt{10}
$$

```{r}
p <- 1/10 # given by problem
q <- 1 - p

no_failure_8_years <-1 - pgeom(7, p) 

no_failure_8_years
```

```{r}
# not failing during the first 8 years
q^8
```

## 2.B 

> What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as an exponential.   

$$
X \sim Exp(1/10) \\
f_X(8) = (0.1) e ^ {-(0.1) 8}, \ x \geq  0 \\
E[X] = \frac{1}{1/10} = 10 \\
\sigma = \sqrt{Var(X)} = \sqrt{\frac{1}{(1/10)^2}} = \sqrt{100} = 10
$$

```{r}
exp_model <- pexp(8, p, lower.tail = FALSE) 

exp_model
```



## 2.C

> What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as a binomial.  (Hint:  0 success in 8 years) 

$$
X \sim Binom(8,1/10) \\ 
P(X = 0) = {8 \choose 0} 0.1^0 (0.9)^8 \\ 
E[X] = np = 8(0.1) = 0.8 \\
\sigma = \sqrt{np(1-p)} = \sqrt{0.72}
$$

```{r}
n <- 8 # n trials
n_success <- 0 # n successes

binom_model <- pbinom(n_success, n, p)

binom_model
```

## 2.D 

> What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as a 
Poisson. 

lamdba is 8/10 since out of the 10 years we want to know the probability it fail after 8 years.

$$
X \sim Pois(8/10) \\
P(X = 0 ) = e^{-0.8}\frac{0.8^0}{0!} = e^{-0.8} \\
E[X] = 0.8 \\
\sigma = \sqrt{0.8} \approx 0.8944 
$$

```{r}
pois_model <- ppois(0, 8/10)

pois_model
```

